---
title: "DTS Analysis Update"
date: "2021-11-09 <br> Updated: `r Sys.Date()`"
---

# ⭐️Overview

On 2021-11-08, we decided to update the tables used in the paper. The analyses run in analysis_01_report.Rmd were overkill for that purpose.

# 📦Load packages

```{r message=FALSE}
library(dplyr, warn.conflicts = FALSE)
library(haven)
library(readr, warn.conflicts = FALSE)
library(tidyr)
library(purrr, warn.conflicts = FALSE)
library(stringr, warn.conflicts = FALSE)
library(officer, warn.conflicts = FALSE)
library(flextable, warn.conflicts = FALSE)
library(broom)
```

# 🌎Connect to UTH server 

```{bash eval=FALSE}
# Don't drill all the way down to live documents because not all of the data is in live documents.
open 'smb://islgpcifs.uthouston.edu/sph_research/Link2Care/'
```

# 📥Import data 

The dts_paper.sav data is created in data_02_clean.Rmd

```{r}
dts_df <- read_rds("/Volumes/Link2Care/Statistical Projects/Robison - Distress Tolerance Scale Paper/dts_paper.rds")
```

```{r}
dim(dts_df) # 298  40
```


# 📈Analysis

## Table 1

### Continuous variables

Calculate the mean and SD for all continuous variables in table 1.

```{r}
table_1_cont <- dts_df %>% 
  # Calculate mean and sd for continuous variables
  summarise(
    across(
      .cols = c(
        age, mms_total, homeless_current_total, phq_8_total, gad_7_total_v1,
        aggression_total, uls_v2_total, dts_total_v2, dts_tolerance_v2,
        dts_appraisal_v2, dts_regulation_v2, dts_absorption_v2
      ),
      .fns = function(x) {
        mean <- mean(x, na.rm = TRUE) %>% round(1) %>% format(nsmall = 1)
        sd <- sd(x, na.rm = TRUE) %>% round(1) %>% format(nsmall = 1)
        out <- paste0(mean, " (", sd, ")")
        out
      }
    )
  ) %>% 
  # Flip vertically
  pivot_longer(
   cols = everything(),
   names_to = "var",
   values_to = "stats"
  )
```


### Categorical variables

Calculate n and percent for all categorical variables in table 1. See R for Epi.

```{r}
table_1_cat <- map_dfr(
  .x = quos(
    gender_v1_f, race_eth_4_cat_f, high_school_grad_f, homeless_mental_treatment_f,
    phq_8_gt_10_f
  ),
  .f = function(x, data = dts_df) {
    data %>% 
      # Drop rows with missing - complete case analysis
      filter(!is.na({{ x }})) %>% 
      # Calculate frequency counts
      count({{ x }}) %>% 
      # Calculate percentages and format as n (%)
      mutate(
        percent = round(n / sum(n) * 100, 1) %>% format(nsmall = 1),
        stats = paste0(n, " (", percent, ")"),
        # Remove white space between opening parenthesis and percentage when 
        # percentage is single digit.
        stats = str_replace_all(stats, "\\( ", "\\(")
      ) %>% 
      mutate(var = names(.)[1]) %>%
      rename(cat = 1) %>%
      select(var, cat, stats)
  }
)
```


## DTS subscale analysis

2021-04-27, From Michael:

Hi Brad. Sorry I did not include you on the Zoom invite. I met with the rest of the team and walked through the analyses. There should be 5 models, one with each DV. All 4 of the Distress Tolerance subscales should be entered at once (with covariates) predicting DVs…

### Create var lists

```{r}
out_vars <- c(
  "PHQ_dep_dichot_total", "aggression_total", "tcu_hs_v2_total", "gad_7_total_v1", 
  "uls_v2_total"
)

pred_vars <- c(
  "dts_tolerance_v2", "dts_absorption_v2", "dts_appraisal_v2", 
  "dts_regulation_v2"
)

# Put predictor variables in a form that is easy to convert into a formula for modeling
pred_vars_mod <- paste(pred_vars, collapse = " + ")

confounders <- c(
  "age", "gender_v1_f", "race_eth_4_cat_f", "high_school_grad_f", 
  "homeless_current_total"
)

# Put confounders in a form that is easy to convert into a formula for modeling
confounders_mod <- paste(confounders, collapse = " + ")

# Combine predictor variables and confounders to form the right-hand side of 
# the model formula
rhs <- paste(pred_vars_mod, confounders_mod, sep = " + ")
```

### Helper function: 

```{r}
tidy_lm <- function(.y, .rhs, .data) {
  # Create a formula for the model function
  f <- as.formula(paste(.y, .rhs, sep = " ~ "))
  # Fit the model
  m <- lm(f, data = .data)
  # Tidy the output
  m %>%
    broom::tidy() %>% 
    # Just keep the coefficients for the dts subscales
    filter(str_detect(term, "dts")) %>% 
    # Make it clear which outcome is being modeled
    mutate(outcome = .y) %>% 
    # Round estimates and p values
    mutate(across(estimate:p.value, ~round(.x, 4))) %>% 
    # Keep and reorder columns of interest
    select(outcome, term, estimate, p.value)
}

# For testing
# tidy_lm(out_vars[1], rhs, dts_df)
```

### Run unadjusted models

2021-07-14, Jordan Neil: For Jillian’s paper, can you send me the output for the 4 DTS models? I don’t use R, but can you create an R markdown file with output that includes the full model diagnostics (e.g., R2, adjusted and unadjusted estimates for DTS and covariates), I can write that up in-text.

```{r}
# For testing
# lm(PHQ_dep_dichot_total ~ dts_tolerance_v2 + dts_absorption_v2 + dts_appraisal_v2 + dts_regulation_v2, data = dts_df) 
```

```{r}
# For testing
# tidy_lm("PHQ_dep_dichot_total", pred_vars_mod, dts_df)
```

```{r}
models_unadj <- map_df(
  .x = out_vars,
  .f = ~ tidy_lm(.x, pred_vars_mod, dts_df)
)
```

```{r}
model_diagnostics_unadj <- map_df(
  .x = out_vars,
  .f = function(x) {
    # Create a formula for the model function
    f <- as.formula(paste(x, pred_vars_mod, sep = " ~ "))
    # Fit the model
    m <- lm(f, data = dts_df)
    # Create data frame of model diagnostics 
    d <- broom::glance(m) %>% 
      # Add the outcome var to the table of output
      mutate(outcome = x) %>% 
      select(outcome, everything()) %>% 
      # Round estimates and p values
      mutate(across(c(r.squared:p.value, logLik:deviance), ~round(.x, 4)))
    d
  }
)
```

### Run Adjusted models

```{r}
# For testing
# lm(PHQ_dep_dichot_total ~ dts_tolerance_v2 + dts_absorption_v2 + dts_appraisal_v2 + dts_regulation_v2 + age + gender_v1_f + race_eth_4_cat_f + high_school_grad_f + homeless_current_total, dts_df)
```

```{r}
# For testing
# tidy_lm("PHQ_dep_dichot_total", rhs, dts_df)
```

```{r}
models_adj <- map_df(
  .x = out_vars,
  .f = ~ tidy_lm(.x, rhs, dts_df)
)
```

```{r}
model_diagnostics_adj <- map_df(
  .x = out_vars,
  .f = function(x) {
    # Create a formula for the model function
    f <- as.formula(paste(x, rhs, sep = " ~ "))
    # Fit the model
    m <- lm(f, data = dts_df)
    # Create data frame of model diagnostics 
    d <- broom::glance(m) %>% 
      # Add the outcome var to the table of output
      mutate(outcome = x) %>% 
      select(outcome, everything()) %>% 
      # Round estimates and p values
      mutate(across(c(r.squared:p.value, logLik:deviance), ~round(.x, 4)))
    d
  }
)
```

All models adjusted for age, sex, race/ethnicity, education (<HS/HS+), and Length of current homelessness.

### Make variable names easier for end users to read

```{r}
models_unadj <- models_unadj %>% 
  mutate(
    term = case_when(
      term == "dts_tolerance_v2" ~ "DTS: Tolerance",
      term == "dts_absorption_v2" ~ "DTS: Absorption",
      term == "dts_appraisal_v2" ~ "DTS: Appraisal",
      term == "dts_regulation_v2" ~ "DTS: Regulation"
    )
  ) %>%
  rename(Predictor = term, Estimate = estimate, p = p.value)
```


```{r}
models_adj <- models_adj %>% 
  mutate(
    term = case_when(
      term == "dts_tolerance_v2" ~ "DTS: Tolerance",
      term == "dts_absorption_v2" ~ "DTS: Absorption",
      term == "dts_appraisal_v2" ~ "DTS: Appraisal",
      term == "dts_regulation_v2" ~ "DTS: Regulation"
    )
  ) %>%
  rename(Predictor = term, Estimate = estimate, p = p.value)
```

### Join adjusted and unadjusted models

Make it so the unadjusted estimate and p-value are placed next the the adjusted estimate and p-value in the table.

```{r}
models_unadj_and_adj <- models_unadj %>% 
  left_join(models_adj, by = c("outcome", "Predictor")) %>% 
  rename(
    `Unadjusted Coef` = Estimate.x, `Unadjusted p` = p.x,
    `Adjusted Coef` = Estimate.y, `Adjusted p` = p.y
  )
```


### Helper function: make_ft

```{r}
make_ft <- function(.y, .data) {
  .data %>% 
    filter(outcome == .y) %>% 
    select(-outcome) %>%
    flextable() %>% 
    set_table_properties(layout = "autofit") %>% 
    # Change font to times new roman
    font(fontname = "Times New Roman", part = "all") %>% 
    # Conditionally format p < 0.05 to green
    color(i = ~ `Unadjusted p` < 0.05, j = 3, color = "#458B00") %>% 
    color(i = ~ `Adjusted p` < 0.05, j = 5, color = "#458B00")
}

# For testing
# make_ft("PHQ_dep_dichot_total", models_unadj_and_adj)
```

```{r}
make_ft_diag <- function(.y, .data) {
  .data %>% 
    filter(outcome == .y) %>% 
    select(-outcome) %>%
    flextable() %>% 
    set_table_properties(layout = "autofit") %>% 
    # Change font to times new roman
    font(fontname = "Times New Roman", part = "all") 
}

# For testing
# make_ft_diag("PHQ_dep_dichot_total", model_diagnostics_unadj)
```

# 📝Load analysis report template for officer

```{r}
dts_report <- read_docx("docs/template_dts_tables.docx") %>% 
  body_replace_text_at_bkm("date", as.character(Sys.Date()))
```

### Create flextables

And add to Word bookmarks

#### Table 1

Continuous

```{r}
flextable(table_1_cont) %>% 
  set_table_properties(layout = "autofit") %>% 
  # Change font to times new roman
  font(fontname = "Times New Roman", part = "all") %>% 
  # Make it easier to read
  
```


```{r}
# Models
walk2(
  .x = out_vars,
  # Bookmark names for Word document
  .y = c(
    "table_depression", "table_aggression", "table_hostility", "table_anxiety",
    "table_uls"
  ),
  .f = function(x, y) {
    ft <- make_ft(x, models_unadj_and_adj)
    out <- get("dts_sub_report", envir = .GlobalEnv)
    out <- out %>% 
      body_replace_flextable_at_bkm(y, ft)
    assign("dts_sub_report", out, envir = .GlobalEnv)
  }
)
```

```{r}
# Unadjusted model diagnostics tables
walk2(
  .x = out_vars,
  # Bookmark names for Word document
  .y = c(
    "table_depression_diag_unadj", "table_aggression_diag_unadj", 
    "table_hostility_diag_unadj", "table_anxiety_diag_unadj", 
    "table_uls_diag_unadj"
  ),
  .f = function(x, y) {
    ft <- make_ft_diag(x, model_diagnostics_unadj)
    out <- get("dts_sub_report", envir = .GlobalEnv)
    out <- out %>% 
      body_replace_flextable_at_bkm(y, ft)
    assign("dts_sub_report", out, envir = .GlobalEnv)
  }
)
```

```{r}
# Adjusted model diagnostics tables
walk2(
  .x = out_vars,
  # Bookmark names for Word document
  .y = c(
    "table_depression_diag_adj", "table_aggression_diag_adj", 
    "table_hostility_diag_adj", "table_anxiety_diag_adj", 
    "table_uls_diag_adj"
  ),
  .f = function(x, y) {
    ft <- make_ft_diag(x, model_diagnostics_adj)
    out <- get("dts_sub_report", envir = .GlobalEnv)
    out <- out %>% 
      body_replace_flextable_at_bkm(y, ft)
    assign("dts_sub_report", out, envir = .GlobalEnv)
  }
)
```

# 📌Generate Word reports

```{r}
print(
  dts_report, 
  "docs/DTS Analysis Report.docx"
)
```

```{r}
print(
  dts_sub_report, 
  "docs/DTS Subscale Analysis Report.docx"
)
```




