---
title: "DTS Analysis Update"
date: "2021-11-09 <br> Updated: `r Sys.Date()`"
---

# ⭐️Overview

On 2021-11-08, we decided to update the tables used in the paper. The analyses run in analysis_01_report.Rmd were overkill for that purpose.

# 📦Load packages

```{r message=FALSE}
library(dplyr, warn.conflicts = FALSE)
library(haven)
library(readr, warn.conflicts = FALSE)
library(tidyr)
library(purrr, warn.conflicts = FALSE)
library(stringr, warn.conflicts = FALSE)
library(officer, warn.conflicts = FALSE)
library(flextable, warn.conflicts = FALSE)
library(broom)
library(ggplot2, warn.conflicts = FALSE)
```

# 🌎Connect to UTH server 

```{bash eval=FALSE}
# Don't drill all the way down to live documents because not all of the data is in live documents.
open 'smb://islgpcifs.uthouston.edu/sph_research/Link2Care/'
```

# 📥Import data 

The dts_paper.sav data is created in data_02_clean.Rmd

```{r}
dts_df <- read_rds("/Volumes/Link2Care/Statistical Projects/Robison - Distress Tolerance Scale Paper/dts_paper.rds")
```

```{r}
dim(dts_df) # 298  40
```


# 📈Analysis

## Table 1

### Continuous variables

Calculate the mean and SD for all continuous variables in table 1.

```{r}
table_1_cont <- dts_df %>% 
  # Calculate mean and sd for continuous variables
  summarise(
    across(
      .cols = c(
        age, mms_total, homeless_current_total, phq_8_total, gad_7_total_v1,
        aggression_total, uls_v2_total, dts_total_v2, dts_tolerance_v2,
        dts_appraisal_v2, dts_regulation_v2, dts_absorption_v2
      ),
      .fns = function(x) {
        mean <- mean(x, na.rm = TRUE) %>% round(1) %>% format(nsmall = 1)
        sd <- sd(x, na.rm = TRUE) %>% round(1) %>% format(nsmall = 1)
        out <- paste0(mean, " (", sd, ")")
        out
      }
    )
  ) %>% 
  # Flip vertically
  pivot_longer(
   cols = everything(),
   names_to = "var",
   values_to = "stats"
  )
```


### Categorical variables

Calculate n and percent for all categorical variables in table 1. See R for Epi.

```{r}
table_1_cat <- map_dfr(
  .x = quos(
    gender_v1_f, race_eth_4_cat_f, high_school_grad_f, homeless_mental_treatment_f,
    phq_8_gt_10_f
  ),
  .f = function(x, data = dts_df) {
    data %>% 
      # Drop rows with missing - complete case analysis
      filter(!is.na({{ x }})) %>% 
      # Calculate frequency counts
      count({{ x }}) %>% 
      # Calculate percentages and format as n (%)
      mutate(
        percent = round(n / sum(n) * 100, 1) %>% format(nsmall = 1),
        stats = paste0(n, " (", percent, ")"),
        # Remove white space between opening parenthesis and percentage when 
        # percentage is single digit.
        stats = str_replace_all(stats, "\\( ", "\\(")
      ) %>% 
      mutate(var = names(.)[1]) %>%
      rename(cat = 1) %>%
      select(var, cat, stats)
  }
)
```


## DTS subscale analysis

2021-04-27, From Michael:

Hi Brad. Sorry I did not include you on the Zoom invite. I met with the rest of the team and walked through the analyses. There should be 5 models, one with each DV. All 4 of the Distress Tolerance subscales should be entered at once (with covariates) predicting DVs…

### Create var lists

```{r}
out_vars <- c(
  "PHQ_dep_dichot_total", "aggression_total", "tcu_hs_v2_total", "gad_7_total_v1", 
  "uls_v2_total"
)

pred_vars <- c(
  "dts_tolerance_v2", "dts_absorption_v2", "dts_appraisal_v2", 
  "dts_regulation_v2"
)

# Put predictor variables in a form that is easy to convert into a formula for modeling
pred_vars_mod <- paste(pred_vars, collapse = " + ")

confounders <- c(
  "age", "gender_v1_f", "race_eth_4_cat_f", "high_school_grad_f", 
  "homeless_current_total"
)

# Put confounders in a form that is easy to convert into a formula for modeling
confounders_mod <- paste(confounders, collapse = " + ")

# Combine predictor variables and confounders to form the right-hand side of 
# the model formula
rhs <- paste(pred_vars_mod, confounders_mod, sep = " + ")
```

### Helper function: 

```{r}
tidy_lm <- function(.y, .rhs, .data) {
  # Create a formula for the model function
  f <- as.formula(paste(.y, .rhs, sep = " ~ "))
  # Fit the model
  m <- lm(f, data = .data)
  # Tidy the output
  m %>%
    broom::tidy() %>% 
    # Just keep the coefficients for the dts subscales
    filter(str_detect(term, "dts")) %>% 
    # Make it clear which outcome is being modeled
    mutate(outcome = .y) %>% 
    # Round estimates and p values
    mutate(across(estimate:p.value, ~round(.x, 4))) %>% 
    # Keep and reorder columns of interest
    select(outcome, term, estimate, p.value)
}

# For testing
# tidy_lm(out_vars[1], rhs, dts_df)
```

### Run unadjusted models

2021-07-14, Jordan Neil: For Jillian’s paper, can you send me the output for the 4 DTS models? I don’t use R, but can you create an R markdown file with output that includes the full model diagnostics (e.g., R2, adjusted and unadjusted estimates for DTS and covariates), I can write that up in-text.

```{r}
# For testing
# lm(PHQ_dep_dichot_total ~ dts_tolerance_v2 + dts_absorption_v2 + dts_appraisal_v2 + dts_regulation_v2, data = dts_df) 
```

```{r}
# For testing
# tidy_lm("PHQ_dep_dichot_total", pred_vars_mod, dts_df)
```

```{r}
models_unadj <- map_df(
  .x = out_vars,
  .f = ~ tidy_lm(.x, pred_vars_mod, dts_df)
)
```

```{r}
model_diagnostics_unadj <- map_df(
  .x = out_vars,
  .f = function(x) {
    # Create a formula for the model function
    f <- as.formula(paste(x, pred_vars_mod, sep = " ~ "))
    # Fit the model
    m <- lm(f, data = dts_df)
    # Create data frame of model diagnostics 
    d <- broom::glance(m) %>% 
      # Add the outcome var to the table of output
      mutate(outcome = x) %>% 
      select(outcome, everything()) %>% 
      # Round estimates and p values
      mutate(across(c(r.squared:p.value, logLik:deviance), ~round(.x, 4)))
    d
  }
)
```

### Run Adjusted models

```{r}
# For testing
# lm(PHQ_dep_dichot_total ~ dts_tolerance_v2 + dts_absorption_v2 + dts_appraisal_v2 + dts_regulation_v2 + age + gender_v1_f + race_eth_4_cat_f + high_school_grad_f + homeless_current_total, dts_df)
```

```{r}
# For testing
# tidy_lm("PHQ_dep_dichot_total", rhs, dts_df)
```

```{r}
models_adj <- map_df(
  .x = out_vars,
  .f = ~ tidy_lm(.x, rhs, dts_df)
)
```

```{r}
model_diagnostics_adj <- map_df(
  .x = out_vars,
  .f = function(x) {
    # Create a formula for the model function
    f <- as.formula(paste(x, rhs, sep = " ~ "))
    # Fit the model
    m <- lm(f, data = dts_df)
    # Create data frame of model diagnostics 
    d <- broom::glance(m) %>% 
      # Add the outcome var to the table of output
      mutate(outcome = x) %>% 
      select(outcome, everything()) %>% 
      # Round estimates and p values
      mutate(across(c(r.squared:p.value, logLik:deviance), ~round(.x, 4)))
    d
  }
)
```

All models adjusted for age, sex, race/ethnicity, education (<HS/HS+), and Length of current homelessness.

### Make variable names easier for end users to read

```{r}
models_unadj <- models_unadj %>% 
  mutate(
    term = case_when(
      term == "dts_tolerance_v2" ~ "DTS: Tolerance",
      term == "dts_absorption_v2" ~ "DTS: Absorption",
      term == "dts_appraisal_v2" ~ "DTS: Appraisal",
      term == "dts_regulation_v2" ~ "DTS: Regulation"
    )
  ) %>%
  rename(Predictor = term, Estimate = estimate, p = p.value)
```


```{r}
models_adj <- models_adj %>% 
  mutate(
    term = case_when(
      term == "dts_tolerance_v2" ~ "DTS: Tolerance",
      term == "dts_absorption_v2" ~ "DTS: Absorption",
      term == "dts_appraisal_v2" ~ "DTS: Appraisal",
      term == "dts_regulation_v2" ~ "DTS: Regulation"
    )
  ) %>%
  rename(Predictor = term, Estimate = estimate, p = p.value)
```

### Join adjusted and unadjusted models

Make it so the unadjusted estimate and p-value are placed next the the adjusted estimate and p-value in the table.

```{r}
models_unadj_and_adj <- models_unadj %>% 
  left_join(models_adj, by = c("outcome", "Predictor")) %>% 
  rename(
    `Unadjusted Coef` = Estimate.x, `Unadjusted p` = p.x,
    `Adjusted Coef` = Estimate.y, `Adjusted p` = p.y
  )
```


# 📄 Create Word results document

## Flextable helper functions

```{r}
make_ft <- function(.y, .data) {
  .data %>% 
    filter(outcome == .y) %>% 
    select(-outcome) %>%
    flextable() %>% 
    # Column width: Trial and error
    # Make a table and play with properties
    width(width = c(1.51, 1.48, 1.2, 1.3, 1.02)) %>% 
    # Change font to times new roman
    font(fontname = "Times New Roman", part = "all") %>% 
    # Conditionally format p < 0.05 to green
    color(i = ~ `Unadjusted p` < 0.05, j = 3, color = "#458B00") %>% 
    color(i = ~ `Adjusted p` < 0.05, j = 5, color = "#458B00")
}

# For testing
# make_ft("PHQ_dep_dichot_total", models_unadj_and_adj)
```

## 📝Load analysis report template for officer

```{r}
dts_report <- read_docx("docs/template_dts_tables.docx") %>% 
  body_replace_text_at_bkm("date", as.character(Sys.Date()))
```

## Create flextables

And add to Word bookmarks

### Table 1

Add overall N

```{r}
dts_report <- dts_report %>% 
  body_replace_text_at_bkm("n_total", as.character(nrow(dts_df)))
```

Continuous vars

```{r}
table_1_cont_ft <- table_1_cont %>% 
  # Make it easier to read
  mutate(var = case_when(
    var == "age" ~ "Age (years)",
    var == "mms_total" ~ "MMSE Total Score",
    var == "homeless_current_total" ~ "Current total time homeless (months)",
    var == "phq_8_total" ~ "PHQ-8 score",
    var == "gad_7_total_v1" ~ "GAD-7 total score",
    var == "aggression_total" ~ "Aggression questionnaire total score",
    var == "uls_v2_total" ~ "Urban Life Stress Scale total score",
    var == "dts_total_v2" ~ "Distress Tolerance Scale total score",
    var == "dts_tolerance_v2" ~ "Tolerance",
    var == "dts_appraisal_v2" ~ "Appraisal",
    var == "dts_regulation_v2" ~ "Regulation",
    var == "dts_absorption_v2" ~ "Absorption"
  )) %>% 
  rename(Variable = var, `Mean (SD)` = stats) %>% 
  flextable() %>%
  # Change font to times new roman
  font(fontname = "Times New Roman", part = "all") %>% 
  # Column width: Trial and error
  # Make a table and play with properties
  width(width = c(4.85, 1.65))
```

Add to flextable

```{r}
dts_report <- dts_report %>% 
  body_replace_flextable_at_bkm("table_1_cont", table_1_cont_ft)
```

Categorical vars

```{r}
table_1_cat_ft <- table_1_cat %>%
  # Make it easier to read
  mutate(var = case_when(
    var == "gender_v1_f" ~ "Gender",
    var == "race_eth_4_cat_f" ~ "Race/Ethnicity",
    var == "high_school_grad_f" ~ "Completed high school education",
    var == "homeless_mental_treatment_f" ~ "Currently receiving treatment for mental health problems",
    var == "phq_8_gt_10_f" ~ "PHQ-8 Score greater than 10"
  )) %>% 
  rename(Variable = var, Category = cat, `n (%)` = stats) %>% 
  flextable() %>%
  # Change font to times new roman
  font(fontname = "Times New Roman", part = "all") %>% 
  # Column width: Trial and error
  # Make a table and play with properties
  width(width = c(3.81, 1.81, 0.88))
```

Add to flextable

```{r}
dts_report <- dts_report %>% 
  body_replace_flextable_at_bkm("table_1_cat", table_1_cat_ft)
```

### DTS subscale models

```{r}
# Models
walk2(
  .x = out_vars,
  # Bookmark names for Word document
  .y = c(
    "table_depression", "table_aggression", "table_hostility", "table_anxiety",
    "table_uls"
  ),
  .f = function(x, y) {
    ft <- make_ft(x, models_unadj_and_adj)
    out <- get("dts_report", envir = .GlobalEnv)
    out <- out %>% 
      body_replace_flextable_at_bkm(y, ft)
    assign("dts_report", out, envir = .GlobalEnv)
  }
)
```

### Explore outliers

2021-11-09, From Jordan Neil: As the effect for model 5 is significantly in the inverse direction that we had hypothesized, we’re interested in whether you could 1) plot the means for the measure and the regression and 2) ascertain whether this sig association is spurious because of some extreme responses?

#### Tolerance 

```{r}
ggplot(dts_df) +
  geom_boxplot(aes(y = uls_v2_total)) +
  theme_bw()
```

#### Tolerance 

```{r}
ggplot(dts_df, aes(y = uls_v2_total, x = dts_tolerance_v2)) +
  geom_point() +
  geom_smooth(method = "lm", formula = 'y ~ x') + 
  labs(
    title = "Plot ULS Total by DTS Tolerance Subscale",
    subtitle = "With regression line"
  ) +
  theme_bw()

ggsave("img/tolerance_plot.png", width = 7.29, height = 4.51)
```

```{r}
dts_report <- dts_report %>% 
  body_replace_img_at_bkm(
    "tolerance_plot", 
    external_img("img/tolerance_plot.png", width = 7, height = 4)
  )
```

#### Absorption 

```{r}
ggplot(dts_df, aes(y = uls_v2_total, x = dts_absorption_v2)) +
  geom_point() +
  geom_smooth(method = "lm", formula = 'y ~ x') + 
  labs(
    title = "Plot ULS Total by DTS Absorption Subscale",
    subtitle = "With regression line"
  ) +
  theme_bw()

ggsave("img/absorption_plot.png", width = 7.29, height = 4.51)
```

```{r}
dts_report <- dts_report %>% 
  body_replace_img_at_bkm(
    "absorption_plot", 
    external_img("img/absorption_plot.png", width = 7, height = 4)
  )
```

#### Appraisal 

```{r}
ggplot(dts_df, aes(y = uls_v2_total, x = dts_appraisal_v2)) +
  geom_point() +
  geom_smooth(method = "lm", formula = 'y ~ x') + 
  labs(
    title = "Plot ULS Total by DTS Appraisal Subscale",
    subtitle = "With regression line"
  ) +
  theme_bw()

ggsave("img/appraisal_plot.png", width = 7.29, height = 4.51)
```

```{r}
dts_report <- dts_report %>% 
  body_replace_img_at_bkm(
    "appraisal_plot", 
    external_img("img/appraisal_plot.png", width = 7, height = 4)
  )
```

#### Regulation 

```{r}
ggplot(dts_df, aes(y = uls_v2_total, x = dts_regulation_v2)) +
  geom_point() +
  geom_smooth(method = "lm", formula = 'y ~ x') + 
  labs(
    title = "Plot ULS Total by DTS Regulation Subscale",
    subtitle = "With regression line"
  ) +
  theme_bw()

ggsave("img/regulation_plot.png", width = 7.29, height = 4.51)
```

```{r}
dts_report <- dts_report %>% 
  body_replace_img_at_bkm(
    "regulation_plot", 
    external_img("img/regulation_plot.png", width = 7, height = 4)
  )
```


# 📌Generate Word reports

```{r}
print(
  dts_report, 
  "docs/DTS Table Updates.docx"
)
```




